Hierarchical Principal Manifold Estimation

### 2023-08-15 paper/main.tex
add backfitting algorithm

\[%
  \mathbf{x}_{jm} = \psi(\mathbf{y}_{jm}) = f_p(\mathbf{y}_{jm}, t_{j}) + f_{g(j)}(\mathbf{y}_{jm}, t_{j}) + f_{i(j)}(\mathbf{y}_{jm}, t_{j}) + f_j(\mathbf{y}_{jm}) + \epsilon_{jm}
.\]%

@@

The smoothing spline functions $f(\mathbf{y})$ minimize 
\[%
  \mathcal{K}_{\lambda}(f) = \sum_{j=1}^{N}\sum_{m=1}^{M_j}w_{jm}\|\mathbf{x}_{jm} - f(\mathbf{y}_{jm})\|^2 + \lambda\|\nabla^{\otimes 2}f\|^2
,\]%
where $\nabla^{\otimes 2}f$ denotes the Hessian matrix of $f$, and $\lambda$ is given. Thus, $f(\mathbf{y})$ takes the form
\begin{equation}
  f_l(\mathbf{y}) = \sum_{j=1}^{N}\sum_{m=1}^{M_j}s_{jm, l} \times \eta_{d}\left(\|\mathbf{y} - \mathbf{y}_{jm}\|\right) + \sum_{p = 1}^{d + 1}\alpha_{p, l} \times \phi_{p}(\mathbf{y}), \quad l = 1, 2, \dots, D, \label{eq:2}
\end{equation}
with $\phi_p(\mathbf{y})$ representing the basis of the linear space of polynomials on $\mathbb{R}^{d}$ with degree $\leq 1$ and $\eta_{d}(y) = y^{4-d}\log y$ when $\log y \neq 0$ and $d$ is even, $\eta_{d}(y) = 0$ when $\log y = 0$ and $d$ is even, and $\eta_{d}(y) = y^{4-d}$ when $d$ is odd.

@@

\begin{equation}
   \left(\begin{array}{ccc}
    2\mathbf{EWE} + 2\lambda\mathbf{E} & 2\mathbf{EWR} & \mathbf{R} \\
    2\mathbf{R}^{T}\mathbf{WE} & 2\mathbf{R}^{T}\mathbf{WR} & \mathbf{0} \\
    \mathbf{R}^{T} & \mathbf{0} & \mathbf{0}
  \end{array}\right)\left(
  \begin{array}{c}
    s_l \\
    \alpha_l \\
    \upsilon_l
  \end{array}
  \right) = \left(
  \begin{array}{c}
    2\mathbf{EW}\mathbf{x}_l \\
    2\mathbf{R}^{T}\mathbf{W}\mathbf{x}_l \\
    \mathbf{0}
  \end{array}
  \right). \label{eq:3}
\end{equation}
The optimal value of $\lambda$ can be found using cross-validation, where with images being randomly split into $K$ folds, functions may be evaluated by the mean squared error, written as
\begin{equation}
  MSE(\lambda) = \frac{1}{K}\sum_{k=1}^{K}\frac{1}{N_u}\sum_{j=1}^{N_u}\frac{1}{M_j}\sum_{m=1}^{M_j}\|\mathbf{x}_{jm} - f_{\lambda}^{(k)}(\mathbf{y}_{jm})\|^2. \label{eq:4}
\end{equation}
Here, $f_{\lambda}^{(k)}(\mathbf{y})$ denotes the estimated spline function given tuning parameter $\lambda$ and with the data for the $k$th fold excluded.

The coefficients and tuning parameter that minimize $\mathcal{K}_{\lambda}(f)$ suffice for estimating $f_{j}(\mathbf{y})$, in which the coefficients are not time-varying. To reach estimates for the varying-coefficient models, equation \ref{eq:3} is solved using the data available at each time point within the relevant study level. To estimate $f_p(\mathbf{y}, t)$, we use $\left\{\mathbf{x}_t\right\}_{t=1}^T$ and $\left\{\mathbf{y}_t\right\}_{t=1}^T$, yielding fitted function $\left\{\hat{f}_t(\mathbf{y})\right\}_{t=1}^T$. Using the maximum and minimum values of $\mathbf{y}_t$ in each dimension, a new $d$-dimensional grid of $M' = \text{max}\left(\left\{M_{j}\right\}_{j = 1}^{N}\right)$ knots, $\left\{\mathbf{y'}_m\right\}_{m=1}^{M'}$, is generated with the knots being evenly spread throughout the parameter space. These parameters serve as common knots that allow the spline coefficients to be compared across time points. They are then used to compute a new set of corresponding outcome values $\left\{\mathbf{x}'_t\right\}_{t=1}^{T}$. The spline functions are reestimated as $\left\{\hat{f}'_t\right\}_{t=1}^{T}$ using $\left\{\mathbf{x}'_t\right\}_{t=1}^{T}$ and $\mathbf{y}'$. These newly generated knots are weighted equally, so an unweighted smoothing spline model is used. Following \cite{greenSilverman1994}, we let

@@

Then the coefficients that minimize $\left(\mathbf{x}'_t - \mathbf{E}'\mathbf{s}_t - \mathbf{R}'^{T}\mathbf{\alpha}_t\right)^{T}\left(\mathbf{x}'_t - \mathbf{E}'\mathbf{s}_t - \mathbf{R}'^{T}\mathbf{\alpha}_t\right) + \gamma\mathbf{s}_t^{T}\mathbf{E}'\mathbf{s}_t$ can be found by solving 
\begin{equation}
  \left(
  \begin{array}{cc}
    \mathbf{E}' + \lambda_{t}\mathbf{I} & \mathbf{R}'^{T} \\
    \mathbf{R}' & \mathbf{0}
  \end{array}
  \right)\left(
  \begin{array}{c}
    \mathbf{s}'_t \\
    \mathbf{\alpha}'_t
  \end{array}
  \right) = \left(
  \begin{array}{c}
    \mathbf{x}'_t \\
    \mathbf{0}
  \end{array}
  \right). \label{eq:5}
\end{equation}

@@

\[%
  g_{\gamma}(t) = \sum_{i=1}^{T}\delta_i\|t - t_i\|^{3} + \sum_{p=1}^{2}\tau_p \phi(t)_p
.\]%
We define matrix $\mathbf{A}$ as $\mathbf{A}_{qr} = \|t_q - t_r\|^{3}$, matrix $\mathbf{T}$ as $\mathbf{T}_{pr} = \phi_p(t_r)$, $\mathbf{\delta} = \left(\delta_1, \dots, \delta_T\right)^{T}$, and $\mathbf{\tau} = \left(\tau_1, \tau_2\right)^{T}$. Then, the coefficients of $g_{\gamma}(t)$, given a value of $\gamma$, can be found by solving
\begin{equation}
  \left(
  \begin{array}{cc}
    \mathbf{A} + \gamma\mathbf{I} & \mathbf{T}^{T} \\
    \mathbf{T} & \mathbf{0}
  \end{array}
  \right)\left(
  \begin{array}{c}
    \mathbf{\delta} \\
    \mathbf{\tau}
  \end{array}
  \right) = \left(
  \begin{array}{c}
    \mathbf{\omega} \\
    \mathbf{0}
  \end{array}
  \right). \label{eq:6}
\end{equation}

As with the tuning process for $\lambda$, the optimal value of $\gamma$ can be identified using cross-validation, with either leave-one-out or $k$-folds cross-validation being appropriate depending on the number of time points available. If using leave-one-out cross-validation, the mean squared error can be calculated as
\begin{equation}
  MSE(\gamma) = \frac{1}{T}\sum_{t=1}^{T}\frac{1}{N_t}\sum_{j=1}^{N_t}\frac{1}{M_j}\sum_{m=1}^{M_j}\|\mathbf{x}_{jm} - f_{\gamma}^{(t)}(\mathbf{y}_{jm}, t_j)\|^2. \label{eq:7}
\end{equation}
We then let $\hat{\gamma}$ be the value of $\gamma$ that minimizes $MSE(\gamma)$.

With a proper estimate of the $\hat{\gamma}$, we may denote the coefficients estimated by $g(t)$ as $\Omega(t) = \left(\mathbf{s}_{\hat{\gamma}}(t)^{T}, \mathbf{\alpha}_{\hat{\gamma}}(t)^{T}\right)^{T}$. We use these coefficients in equation \ref{eq:2} in place of the constant coefficients. Therefore, the varying-coefficient models have the following outputs:
\begin{equation}
  f(\mathbf{y}, t) = \sum_{m=1}^{M'}\mathbf{s}_{\hat{\gamma}, l}(t) \eta_{d}\left(\|\mathbf{y} - \mathbf{y'}_m\|\right) + \sum_{p=1}^{d + 1}\mathbf{\alpha}_{\hat{\gamma}, p, l}(t) \phi_p(\mathbf{y}), \quad l = 1, \dots, D. \label{eq:8}
\end{equation}
The fitting process for the varying-coefficient smoothing spline model is given formally as Algorithm \ref{alg:varying_spline}.

@@

\begin{algorithm}
\caption{Varying-Coefficient Smoothing Spline Model}\label{alg:varying_spline}
  \KwData{$D$-dimensional output values $\left\{\mathbf{x}_{jm}\right\}_{j = 1, m = 1}^{N, M_{j}}$, $d$-dimensional input values $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, weights $\left\{w_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, observation times $\left\{t_{j}\right\}_{j=1}^{N}$, candidate tuning parameters $\left\{\lambda_l\right\}_{l=1}^{L}$, $\left\{\gamma_p\right\}_{p=1}^{P}$, and number of cross-validation folds $K$.}
  \KwResult{Analytic formula of $\hat{f}(\mathbf{y}, t)$ and optimal tuning parameter $\hat{\gamma}$.}

\For{$t = 1, 2, \dots, T$} {
  Set $U_t$ equal to the number of images included in the data at time $t$\;
  \For{$l = 1, 2, \dots, L$} {
    Solve equation \ref{eq:3} to estimate $f_{\lambda_l, t}(\mathbf{y})$ with inputs $\mathbf{x}_t$, $\mathbf{y}_t$, and $w_t$\;
    \For{$u = 1, 2, \dots, U_t$} {
      Solve equation \ref{eq:3} to estimate $f^{(-u)}_{\lambda_l, t}(\mathbf{y})$ with inputs $\mathbf{x}^{(-u)}_t$, $\mathbf{y}^{(-u)}_t$, $w^{(-u)}_t$\;
      Estimate $MSE(\lambda_{l})_u$ using equation \ref{eq:4}\;
    }
    Set $MSE(\lambda_l) = \frac{1}{U}\sum_{u=1}^{U}MSE(\lambda_l)_u$\;
  }
  $\hat{\lambda}_t = \arg \min_{\lambda} MSE(\lambda)$\;
  $f_t(\mathbf{y}) = f_{\hat{\lambda}_t}(\mathbf{y})$\;
}

  Let $M' = \max(M_{j})$, $\left\{\mathbf{y}'_{m}\right\}_{m=1}^{M'}$ be a grid spanning the range of values of $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N, M_j}$\;
  \For{$t = 1, 2, \dots, T$} {
    Compute $\left\{\mathbf{x'}_{m, t}\right\}_{m=1}^{M'} = \left\{f_t(\mathbf{y'}_m)\right\}_{m=1}^{M'}$\;
    Solve equation \ref{eq:5} to compute $\mathbf{s'}_t$ and $\mathbf{\alpha'}_t$\;
    Set $\mathbf{\omega}_t = \left[\mathbf{s'}_t^{T}, \mathbf{\alpha'}_t^{T}\right]$\;
  }
  \For{$p = 1, 2, \dots, P$} {
    Compute $g_{\gamma_p}(t)$ by solving equation \ref{eq:6}\;
    \For{$t = 1, 2, \dots, T$} {
      Compute $g_{\gamma_p}^{(t)}$ by solving \ref{eq:6}\;
    }
    Estimate $MSE(\gamma_p)$ using \ref{eq:7}\;
  }
  $\hat{\gamma} = \arg \min_{\gamma} MSE(\gamma)$\;
  $\hat{f}(\mathbf{y}, t) = f_{\hat{\gamma}}(\mathbf{y}, t)$, where the form of $\hat{f}(\mathbf{y}, t)$ is given in \ref{eq:8}.
\end{algorithm}

@@

As discussed in \cite{hastieGeneralizedAdditiveModels1990}, the additive model $\psi(\mathbf{y})$ can be estimated using a backfitting procedure, which uses the model's additive nature to iteratively fit each distinct smoothing spline function until convergence. Specifically, we use the fact that, under the model,

\[%
f_p(\mathbf{y}_{ijk}, t_{ij}) = E\left[\mathbf{x}_{ijk} - f_{g(i)}(\mathbf{y}_{ijk}, t_{ij}) - f_{i}(\mathbf{y}_{ijk}, t_{ij}) - f_{ij}(\mathbf{y}_{ijk})\right]
,\]%
with a similar result holding for each of the smooth functions making up the additive model. Thus, the backfitting algorithm to fit the proposed model is given below.

\zielinski{The following algorithm is currently used as a placeholder while I work out the specifics.}

\zielinski{Try using a ratio of the norm of the difference in model coefficients and the norm of the old model to measure convergence. This will set the change in the model coefficients (as measured by the norm of the difference vector/matrix) relative to the magnitude of the model coefficients.}

\begin{algorithm}
\caption{Backfitting Algorithm for Additive Spline Model}\label{alg:backfitting}
  \KwData{$D$-dimensional output values $\left\{\mathbf{x}_{ijk}\right\}_{i=1, j=1, k=1}^{N, n_i, m_{ij}}$, $d$-dimensional input values $\left\{\mathbf{y}_{ijk}\right\}_{i=1, j=1, k=1}^{N, n_i, m_{ij}}$, and observation times $\left\{t_{ij}\right\}_{i=1, j=1}^{N, n_i}$.}
  \KwResult{Analytic formulas of $\hat{f}_p$, $\left\{\hat{f}_g\right\}_{g=1}^{G}, \left\{\hat{f}_i\right\}_{i=1}^{N}$, and $\left\{\hat{f}_{ij}\right\}_{i=1, j=1}^{N, n_i}$, and optimal tuning parameters $\hat{\lambda}_p$, $\hat{\gamma}_p$, $\left\{\hat{\lambda}_g\right\}_{g=1}^{G}$, $\left\{\hat{\gamma}\right\}_{g=1}^{G}$, $\left\{\hat{\lambda}_{i}\right\}_{i=1}^{N}$, $\left\{\hat{\gamma}_{i}\right\}_{i=1}^{N}$, $\left\{\hat{\lambda}_{ij}\right\}_{i=1, j=1}^{N, n_i}$.}

>>>>

\begin{equation}
  \mathbf{x}_{jm} = \psi(\mathbf{y}_{jm}) = f_p(\mathbf{y}_{jm}, t_{j}) + f_{g(j)}(\mathbf{y}_{jm}, t_{j}) + f_{i(j)}(\mathbf{y}_{jm}, t_{j}) + f_j(\mathbf{y}_{jm}) + \epsilon_{jm}. \label{eq:2}
\end{equation}

@@

The smoothing spline functions $f(\mathbf{y})$ minimize 
\[%
  \mathcal{K}_{\lambda}(f) = \sum_{j=1}^{N}\sum_{m=1}^{M_j}w_{jm}\|\mathbf{x}_{jm} - f(\mathbf{y}_{jm})\|^2 + \lambda\|\nabla^{\otimes 2}f\|^2
,\]%
where $\nabla^{\otimes 2}f$ denotes the Hessian matrix of $f$, and $\lambda$ is given. Thus, $f(\mathbf{y})$ takes the form
\begin{equation}
  f_l(\mathbf{y}) = \sum_{j=1}^{N}\sum_{m=1}^{M_j}s_{jm, l} \times \eta_{d}\left(\|\mathbf{y} - \mathbf{y}_{jm}\|\right) + \sum_{p = 1}^{d + 1}\alpha_{p, l} \times \phi_{p}(\mathbf{y}), \quad l = 1, 2, \dots, D, \label{eq:3}
\end{equation}
with $\phi_p(\mathbf{y})$ representing the basis of the linear space of polynomials on $\mathbb{R}^{d}$ with degree $\leq 1$ and $\eta_{d}(y) = y^{4-d}\log y$ when $\log y \neq 0$ and $d$ is even, $\eta_{d}(y) = 0$ when $\log y = 0$ and $d$ is even, and $\eta_{d}(y) = y^{4-d}$ when $d$ is odd.

@@

\begin{equation}
   \left(\begin{array}{ccc}
    2\mathbf{EWE} + 2\lambda\mathbf{E} & 2\mathbf{EWR} & \mathbf{R} \\
    2\mathbf{R}^{T}\mathbf{WE} & 2\mathbf{R}^{T}\mathbf{WR} & \mathbf{0} \\
    \mathbf{R}^{T} & \mathbf{0} & \mathbf{0}
  \end{array}\right)\left(
  \begin{array}{c}
    s_l \\
    \alpha_l \\
    \upsilon_l
  \end{array}
  \right) = \left(
  \begin{array}{c}
    2\mathbf{EW}\mathbf{x}_l \\
    2\mathbf{R}^{T}\mathbf{W}\mathbf{x}_l \\
    \mathbf{0}
  \end{array}
  \right). \label{eq:4}
\end{equation}
The optimal value of $\lambda$ can be found using cross-validation, where with images being randomly split into $K$ folds, functions may be evaluated by the mean squared error, written as
\begin{equation}
  MSE(\lambda) = \frac{1}{K}\sum_{k=1}^{K}\frac{1}{N_u}\sum_{j=1}^{N_u}\frac{1}{M_j}\sum_{m=1}^{M_j}\|\mathbf{x}_{jm} - f_{\lambda}^{(k)}(\mathbf{y}_{jm})\|^2. \label{eq:5}
\end{equation}
Here, $f_{\lambda}^{(k)}(\mathbf{y})$ denotes the estimated spline function given tuning parameter $\lambda$ and with the data for the $k$th fold excluded.

The coefficients and tuning parameter that minimize $\mathcal{K}_{\lambda}(f)$ suffice for estimating $f_{j}(\mathbf{y})$, in which the coefficients are not time-varying. To reach estimates for the varying-coefficient models, equation \ref{eq:4} is solved using the data available at each time point within the relevant study level. To estimate $f_p(\mathbf{y}, t)$, we use $\left\{\mathbf{x}_t\right\}_{t=1}^T$ and $\left\{\mathbf{y}_t\right\}_{t=1}^T$, yielding fitted function $\left\{\hat{f}_t(\mathbf{y})\right\}_{t=1}^T$. Using the maximum and minimum values of $\mathbf{y}_t$ in each dimension, a new $d$-dimensional grid of $M' = \text{max}\left(\left\{M_{j}\right\}_{j = 1}^{N}\right)$ knots, $\left\{\mathbf{y'}_m\right\}_{m=1}^{M'}$, is generated with the knots being evenly spread throughout the parameter space. These parameters serve as common knots that allow the spline coefficients to be compared across time points. They are then used to compute a new set of corresponding outcome values $\left\{\mathbf{x}'_t\right\}_{t=1}^{T}$. The spline functions are reestimated as $\left\{\hat{f}'_t\right\}_{t=1}^{T}$ using $\left\{\mathbf{x}'_t\right\}_{t=1}^{T}$ and $\mathbf{y}'$. These newly generated knots are weighted equally, so an unweighted smoothing spline model is used. Following \cite{greenSilverman1994}, we let

@@

Then the coefficients that minimize $\left(\mathbf{x}'_t - \mathbf{E}'\mathbf{s}_t - \mathbf{R}'^{T}\mathbf{\alpha}_t\right)^{T}\left(\mathbf{x}'_t - \mathbf{E}'\mathbf{s}_t - \mathbf{R}'^{T}\mathbf{\alpha}_t\right) + \gamma\mathbf{s}_t^{T}\mathbf{E}'\mathbf{s}_t$ can be found by solving 
\begin{equation}
  \left(
  \begin{array}{cc}
    \mathbf{E}' + \lambda_{t}\mathbf{I} & \mathbf{R}'^{T} \\
    \mathbf{R}' & \mathbf{0}
  \end{array}
  \right)\left(
  \begin{array}{c}
    \mathbf{s}'_t \\
    \mathbf{\alpha}'_t
  \end{array}
  \right) = \left(
  \begin{array}{c}
    \mathbf{x}'_t \\
    \mathbf{0}
  \end{array}
  \right). \label{eq:6}
\end{equation}

@@

\[%
  g_{\gamma}(t) = \sum_{i=1}^{T}\delta_i\|t - t_i\|^{3} + \sum_{p=1}^{2}\tau_p \phi(t)_p
.\]%
We define matrix $\mathbf{A}$ as $\mathbf{A}_{qr} = \|t_q - t_r\|^{3}$, matrix $\mathbf{T}$ as $\mathbf{T}_{pr} = \phi_p(t_r)$, $\mathbf{\delta} = \left(\delta_1, \dots, \delta_T\right)^{T}$, and $\mathbf{\tau} = \left(\tau_1, \tau_2\right)^{T}$. Then, the coefficients of $g_{\gamma}(t)$, given a value of $\gamma$, can be found by solving
\begin{equation}
  \left(
  \begin{array}{cc}
    \mathbf{A} + \gamma\mathbf{I} & \mathbf{T}^{T} \\
    \mathbf{T} & \mathbf{0}
  \end{array}
  \right)\left(
  \begin{array}{c}
    \mathbf{\delta} \\
    \mathbf{\tau}
  \end{array}
  \right) = \left(
  \begin{array}{c}
    \mathbf{\omega} \\
    \mathbf{0}
  \end{array}
  \right). \label{eq:7}
\end{equation}

As with the tuning process for $\lambda$, the optimal value of $\gamma$ can be identified using cross-validation, with either leave-one-out or $k$-folds cross-validation being appropriate depending on the number of time points available. If using leave-one-out cross-validation, the mean squared error can be calculated as
\begin{equation}
  MSE(\gamma) = \frac{1}{T}\sum_{t=1}^{T}\frac{1}{N_t}\sum_{j=1}^{N_t}\frac{1}{M_j}\sum_{m=1}^{M_j}\|\mathbf{x}_{jm} - f_{\gamma}^{(t)}(\mathbf{y}_{jm}, t_j)\|^2. \label{eq:8}
\end{equation}
We then let $\hat{\gamma}$ be the value of $\gamma$ that minimizes $MSE(\gamma)$.

With a proper estimate of the $\hat{\gamma}$, we may denote the coefficients of $g(t)$ as $\Omega = \left(\delta^{T}, \tau^{T}\right)$, and those estimated by $g(t)$ as $\omega(t) = \left(\mathbf{s}_{\hat{\gamma}}(t)^{T}, \mathbf{\alpha}_{\hat{\gamma}}(t)^{T}\right)^{T}$. We use these coefficients in equation \ref{eq:3} in place of the constant coefficients. Therefore, the varying-coefficient models have the following outputs:
\begin{equation}
  f(\mathbf{y}, t) = \sum_{m=1}^{M'}\mathbf{s}_{\hat{\gamma}, l}(t) \eta_{d}\left(\|\mathbf{y} - \mathbf{y'}_m\|\right) + \sum_{p=1}^{d + 1}\mathbf{\alpha}_{\hat{\gamma}, p, l}(t) \phi_p(\mathbf{y}), \quad l = 1, \dots, D. \label{eq:9}
\end{equation}
The fitting process for the varying-coefficient smoothing spline model is given formally as Algorithm \ref{alg:varying_spline}.

@@

\begin{algorithm}
\caption{Varying-Coefficient Smoothing Spline Model}\label{alg:varying_spline}
  \KwData{$D$-dimensional output values $\left\{\mathbf{x}_{jm}\right\}_{j = 1, m = 1}^{N, M_{j}}$, $d$-dimensional input values $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, weights $\left\{w_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, observation times $\left\{t_{j}\right\}_{j=1}^{N}$, candidate tuning parameters $\left\{\lambda_l\right\}_{l=1}^{L}$, $\left\{\gamma_p\right\}_{p=1}^{P}$, and number of cross-validation folds $K$.}
  \KwResult{Analytic formula of $\hat{f}(\mathbf{y}, t)$, coefficients $\Omega = \left(\mathbf{\delta}^{T}, \mathbf{\tau}^{T}\right)$, and optimal tuning parameter $\hat{\gamma}$.}

\For{$t = 1, 2, \dots, T$} {
  Set $U_t$ equal to the number of images included in the data at time $t$\;
  \For{$l = 1, 2, \dots, L$} {
    Solve equation \ref{eq:4} to estimate $f_{\lambda_l, t}(\mathbf{y})$ with inputs $\mathbf{x}_t$, $\mathbf{y}_t$, and $w_t$\;
    \For{$u = 1, 2, \dots, U_t$} {
      Solve equation \ref{eq:4} to estimate $f^{(-u)}_{\lambda_l, t}(\mathbf{y})$ with inputs $\mathbf{x}^{(-u)}_t$, $\mathbf{y}^{(-u)}_t$, $w^{(-u)}_t$\;
      Estimate $MSE(\lambda_{l})_u$ using equation \ref{eq:5}\;
    }
    Set $MSE(\lambda_l) = \frac{1}{U}\sum_{u=1}^{U}MSE(\lambda_l)_u$\;
  }
  $\hat{\lambda}_t = \arg \min_{\lambda} MSE(\lambda)$\;
  $f_t(\mathbf{y}) = f_{\hat{\lambda}_t}(\mathbf{y})$\;
}

  Let $M' = \max(M_{j})$, $\left\{\mathbf{y}'_{m}\right\}_{m=1}^{M'}$ be a grid spanning the range of values of $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N, M_j}$\;
  \For{$t = 1, 2, \dots, T$} {
    Compute $\left\{\mathbf{x'}_{m, t}\right\}_{m=1}^{M'} = \left\{f_t(\mathbf{y'}_m)\right\}_{m=1}^{M'}$\;
    Solve equation \ref{eq:6} to compute $\mathbf{s'}_t$ and $\mathbf{\alpha'}_t$\;
    Set $\mathbf{\omega}_t = \left[\mathbf{s'}_t^{T}, \mathbf{\alpha'}_t^{T}\right]$\;
  }
  \For{$p = 1, 2, \dots, P$} {
    Compute $g_{\gamma_p}(t)$ by solving equation \ref{eq:7}\;
    \For{$t = 1, 2, \dots, T$} {
      Compute $g_{\gamma_p}^{(t)}$ by solving \ref{eq:7}\;
    }
    Estimate $MSE(\gamma_p)$ using \ref{eq:8}\;
  }
  $\hat{\gamma} = \arg \min_{\gamma} MSE(\gamma)$\;
  $\hat{f}(\mathbf{y}, t) = f_{\hat{\gamma}}(\mathbf{y}, t)$, where the form of $\hat{f}(\mathbf{y}, t)$ is given in \ref{eq:9}.
\end{algorithm}

While the procedure outlined in Algorithm \ref{alg:varying_spline} yields an appropriate estimator when working with unnested data, this approach must be adapted to work within the additive model structure of model \ref{eq:2}. As discussed in \cite{hastieGeneralizedAdditiveModels1990}, the additive model $\psi(\mathbf{y})$ can be estimated using a backfitting procedure, which uses the model's additive nature to iteratively fit each distinct smoothing spline function until convergence. Specifically, we use the fact that, under the model,

\[%
  f_p(\mathbf{y}_{jm}, t_{j}) = E\left[\mathbf{x}_{jm} - f_{g(j)}(\mathbf{y}_{jm}, t_{j}) - f_{i(j)}(\mathbf{y}_{jm}, t_{j}) - f_{j}(\mathbf{y}_{jm})\right]
,\]%
with a similar result holding for each of the smooth functions making up the additive model. Thus, the backfitting algorithm to fit the proposed model is given below.

\begin{algorithm}
\caption{Backfitting Algorithm for Additive Spline Model}\label{alg:backfitting}
  \KwData{$D$-dimensional output values $\left\{\mathbf{x}_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, $d$-dimensional input values $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N, M_{j}}$, observation times $\left\{t_{j}\right\}_{j=1}^{N}$, group and individual indicator values $\left\{g(j)\right\}_{j=1}^{N}$ and $\left\{i(j)\right\}_{j=1}^{N}$, and $\epsilon$.}
  \KwResult{Analytic formulas of $\hat{f}_p$, $\left\{\hat{f}_g\right\}_{g=1}^{G}, \left\{\hat{f}_i\right\}_{i=1}^{I}$, and $\left\{\hat{f}_{j}\right\}_{j=1}^{N}$.}

  Estimate $f_{p}^{(0)}(\mathbf{y}, t)$ and $\Omega_{p}^{(0)}$ using Algorithm \ref{alg:varying_spline} with inputs $\left\{\mathbf{x}_{jm}\right\}_{j=1, m=1}^{N, M_j}$, $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N, M_j}$, $\left\{w_{jm}\right\}_{j=1, m=1}^{N, M_j}$, $\left\{t_{j}\right\}_{j=1}^{N}$, $\left\{\lambda_l\right\}_{l=1}^{L}$, $\left\{\gamma_p\right\}_{p=1}^{P}$, and $K$\;
  \For{$g = 1, 2, \dots, G$} {
    Estimate $f_{g}^{(0)}(\mathbf{y}, t)$ and $\Omega_{g}^{(0)}$ using Algorithm \ref{alg:varying_spline} with inputs $\left\{\mathbf{x}_{jm} - f_{p}^{(0)}(\mathbf{y}_{jm}, t_j)\right\}_{j=1, m=1}^{N_g, M_j}$, $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N_g, M_j}$, $\left\{w_{jm}\right\}_{j=1, m=1}^{N_g, M_j}$, $\left\{t_j\right\}_{j=1}^{N_g}$, $\left\{\lambda_l\right\}_{l=1}^{L}$, $\left\{\gamma_p\right\}_{p=1}^{P}$, $K$\;
  }
  \For{$i = 1, 2, \dots, I$} {
    Estimate $f_{i}^{(0)}(\mathbf{y}, t)$ and $\Omega_{i}^{(0)}$ using Algorithm \ref{alg:varying_spline} with inputs $\left\{\mathbf{x}_{jm} - f_{p}^{(0)}(\mathbf{y}_{jm}, t_j) - f_{g(j)}^{(0)}(\mathbf{y}_{jm}, t_j)\right\}_{j=1, m=1}^{N_i, M_j}$, $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N_i, M_j}$, $\left\{w_{jm}\right\}_{j=1, m=1}^{N_i, M_j}$, $\left\{t_j\right\}_{j=1}^{N_i}$, $\left\{\lambda_l\right\}_{l=1}^{L}$, $\left\{\gamma_p\right\}_{p=1}^{P}$, and $K$\;
  }
  \For{$j = 1, 2, \dots, N$} {
    Estimate $f_{j}^{(0)}(\mathbf{y})$ and $\omega_{j}^{(0)}$ by solving \ref{eq:4} with inputs $\left\{\mathbf{x}_{jm} - f_{p}^{(0)}(\mathbf{y}_{jm}, t_j) - f_{g(j)}^{(0)}(\mathbf{y}_{jm}, t_j) - f_{i(j)}^{(0)}(\mathbf{y}_{jm}, t_j)\right\}_{m=1}^{M_j}$, $\left\{\mathbf{y}_{jm}\right\}_{m=1}^{M_j}$, $\left\{w_{jm}\right\}_{m = 1}^{M_j}$, $\left\{\lambda_l\right\}_{l=1}^{L}$, $K$\;
  }
  $n \gets 0$\;
  \While{$\max\left(\|\Omega_p^{(n)} - \Omega_{p}^{(n-1)}\|, \left\{\|\Omega_{g}^{(n)} - \Omega_{g}^{(n-1)}\|\right\}_{g=1}^{G}, \left\{\|\Omega_{i}^{(n)} - \Omega_{i}^{(n-1)}\|\right\}_{i=1}^{I}, \left\{\|\omega_{j}^{(n)} - \omega_{j}^{(n-1)}\|\right\}_{j=1}^{N}\right) \geq \epsilon$} {
    Estimate $f_{p}^{(n+1)}(\mathbf{y}, t)$ and $\Omega_{p}^{(n+1)}$ using Algorithm \ref{alg:varying_spline} with inputs $\left\{\mathbf{x}_{jm} - f_{g(j)}^{(n)}(\mathbf{y}_{jm}, t_j) - f_{i(j)}^{(n)}(\mathbf{y}_{jm}, t_j) - f_j^{(n)}(\mathbf{y}_{jm})\right\}_{j=1, m=1}^{N, M_j}$, $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N, M_j}$, $\left\{w_{jm}\right\}_{j=1, m=1}^{N, M_j}$, $\left\{t_{j}\right\}_{j=1}^{N}$, $\left\{\lambda_l\right\}_{l=1}^{L}$, $\left\{\gamma_p\right\}_{p=1}^{P}$, and $K$\;
    \For{$g = 1, 2, \dots, G$} {
      Estimate $f_{g}^{(n+1)}(\mathbf{y}, t)$ and $\Omega_{g}^{(n+1)}$ using Algorithm \ref{alg:varying_spline} with inputs $\left\{\mathbf{x}_{jm} - f_{p}^{(n+1)}(\mathbf{y}_{jm}, t_j) - f_{i(j)}^{(n)}(\mathbf{y}_{jm}, t_j) - f_{j}^{(n)}(\mathbf{y}_{jm})\right\}_{j=1, m=1}^{N_g, M_j}$, $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N_g, M_j}$, $\left\{w_{jm}\right\}_{j=1, m=1}^{N_g, M_j}$, $\left\{t_j\right\}_{j=1}^{N_g}$, $\left\{\lambda_l\right\}_{l=1}^{L}$, $\left\{\gamma_p\right\}_{p=1}^{P}$, $K$\;
    }
    \For{$i = 1, 2, \dots, I$} {
      Estimate $f_{i}^{(n+1)}(\mathbf{y}, t)$ and $\Omega_{i}^{(n+1)}$ using Algorithm \ref{alg:varying_spline} with inputs $\left\{\mathbf{x}_{jm} - f_{p}^{(n+1)}(\mathbf{y}_{jm}, t_j) - f_{g(j)}^{(n+1)}(\mathbf{y}_{jm}, t_j) - f_j^{(n)}(\mathbf{y}_{jm})\right\}_{j=1, m=1}^{N_i, M_j}$, $\left\{\mathbf{y}_{jm}\right\}_{j=1, m=1}^{N_i, M_j}$, $\left\{w_{jm}\right\}_{j=1, m=1}^{N_i, M_j}$, $\left\{t_j\right\}_{j=1}^{N_i}$, $\left\{\lambda_l\right\}_{l=1}^{L}$, $\left\{\gamma_p\right\}_{p=1}^{P}$, and $K$\;
    }
    \For{$j = 1, 2, \dots, N$} {
      Estimate $f_{j}^{(n+1)}(\mathbf{y})$ and $\omega_{j}^{(n+1)}$ by solving \ref{eq:4} with inputs $\left\{\mathbf{x}_{jm} - f_{p}^{(n+1)}(\mathbf{y}_{jm}, t_j) - f_{g(j)}^{(n+1)}(\mathbf{y}_{jm}, t_j) - f_{i(j)}^{(n+1)}(\mathbf{y}_{jm}, t_j)\right\}_{m=1}^{M_j}$, $\left\{\mathbf{y}_{jm}\right\}_{m=1}^{M_j}$, $\left\{w_{jm}\right\}_{m = 1}^{M_j}$, $\left\{\lambda_l\right\}_{l=1}^{L}$, $K$\;
    }
    $n \gets n + 1$
  }
\end{algorithm}

@@
